Location and Scale-Invariant Power Transformations for Transforming Data to

### 2024-10-16 manuscript_rmarkdown/manuscript.Rmd
Added missing doi to references.

```{=latex}
\begin{equation}
\label{eqn:yeo-johnson-original}
\phi_{\text{YJ}}^\lambda (x_i) = 
\begin{cases}
\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
\end{cases}
\end{equation}
```
The $\lambda$-parameter is typically optimised using maximum likelihood estimation under the assumption that the transformed feature is normally distributed.
As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2021-kq].

@@

Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive.
Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:decreased-normality}.
As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*.
To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
Briefly, this process requires identifying outliers in the data and weighting such instances during the optimisation process.
@Raymaekers2021-kq achieve this through weighted maximum likelihood estimation.
However, because this procedure iteratively estimates and updates $\lambda$, it can not be used here to simultaneously estimate $\lambda$, $x_0$ and $s$ for location- and scale-invariant power transformations.
Nonetheless, as a procedure, weighted MLE can be used for estimating the transformation, shift and scale parameters.

Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
Compared to @Raymaekers2021-kq, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
The weighted location- and scale-invariant Box-Cox log-likelihood function is:

@@

- The weighting function uses the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$.
  After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
  Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
  Then $\dot{x}_i = z_i$.

>>>>

```{=latex}
\begin{equation}
\label{eqn:yeo-johnson-original}
\phi_{\text{YJ}}^\lambda (x_i) = 
\begin{cases}
\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
\end{cases}
\end{equation}
```
The $\lambda$-parameter is typically optimised using maximum likelihood estimation under the assumption that the transformed feature is normally distributed.
As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2024-zf].

@@

Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive.
Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:decreased-normality}.
As indicated by @Raymaekers2024-zf, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*.
To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
Briefly, this process requires identifying outliers in the data and weighting such instances during the optimisation process.
@Raymaekers2024-zf achieve this through weighted maximum likelihood estimation.
However, because this procedure iteratively estimates and updates $\lambda$, it can not be used here to simultaneously estimate $\lambda$, $x_0$ and $s$ for location- and scale-invariant power transformations.
Nonetheless, as a procedure, weighted MLE can be used for estimating the transformation, shift and scale parameters.

Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
Compared to @Raymaekers2024-zf, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
The weighted location- and scale-invariant Box-Cox log-likelihood function is:

@@

- The weighting function uses the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$.
  After @Raymaekers2024-zf, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
  Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
  Then $\dot{x}_i = z_i$.
